Secure Composition of Robust and Optimising Compilers

### 2024-10-14 tech/main.tex
tech

\paragraph{Spatial Memory Safety}
\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
For emphasis, we also use a custom trace-model for spatial memory safety.
\begin{definition}{\definitionlabel[\glsfirst{sms} Traces]{smstraces}}
\[
  \begin{array}{rcll}
    \text{(Base-Events)}&\smsEvent &:=& \smsAlloc \mid \smsDealloc \mid \smsUse \\
    \text{(Events)}&\smsEvent &:=& \smsEvent \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]
\end{definition}

@@

    \typerule{mapping-ms-sms-empty}{}{
      \mmap{\msev{\emptyevent}}{\smsev{\emptyevent}}
    }{mapping-ms-sms-empty}
    %
    \typerule{mapping-ms-sms-crash}{}{
      \mmap{\msev{\lightning}}{\smsev{\lightning}}
    }{mapping-ms-sms-crash}
    %
    \typerule{mapping-ms-sms-alloc}{}{
      \mmap{\msAlloc}{\smsAlloc}
    }{mapping-ms-sms-alloc}
    %
    \typerule{mapping-ms-sms-dealloc}{}{
      \mmap{\msDealloc}{\smsDealloc}
    }{mapping-ms-sms-dealloc}
    %
    \typerule{mapping-ms-sms-use}{}{
      \mmap{\msUse}{\smsUse}
    }{mapping-ms-sms-use}
  \end{center}
  \begin{center}
    \judgbox{\mmap{\msTrace}{\smsTrace}}{,,Translate \gls*{ms}-traces to \gls*{sms}-traces.''}

@@

    \typerule{mapping-ms-spec-empty}{}{
      \mmap{\msev{\emptyevent}}{\specev{\emptyevent}}
    }{mapping-ms-spec-empty}
    %
    \typerule{mapping-ms-spec-crash}{}{
      \mmap{\msev{\lightning}}{\specev{\lightning}}
    }{mapping-ms-spec-crash}
    %
    \typerule{mapping-ms-spec-alloc}{}{
      \mmap{\msAlloc}{\specAny[\unlock]}
    }{mapping-ms-spec-alloc}
    %
    \typerule{mapping-ms-spec-dealloc}{}{
      \mmap{\msDealloc}{\specAny[\unlock]}
    }{mapping-ms-spec-dealloc}
    %
    \typerule{mapping-ms-spec-use}{}{
      \mmap{\msUse}{\specAny[\unlock]}
    }{mapping-ms-spec-use}
  \end{center}
  \begin{center}
    \judgbox{\mmap{\msTrace}{\specTrace}}{,,Translate \gls*{ms}-traces to \gls*{spec}-traces.''}

@@

\subsubsection{\gls*{tms} Monitor}\label{subsubsec-mon-tms}
In order to check \gls*{tms} on traces mechanically, we introduce a monitor that works on more abstract monitor-actions, without any other events besides those relevant to temporal memory safety.

@@

    \typerule{tmsmon-refl}{
      \tmsMonitor[.A]=\emptyset
    }{
      \monitorcheck{\tmsMonitor}{\tmsMonitor}{\hole{\cdot}}
    }{tmsmon-refl}
    %
    \typerule{tmsmon-trans}{
      \monitorcheck[]{\tmsMonitor}{\tmsMonitor[']}{\tmsEvent}&
      \monitorcheck{\tmsMonitor[']}{\tmsMonitor['']}{\tmsTrace}&
    }{
      \monitorcheck{\tmsMonitor}{\tmsMonitor['']}{\tmsEvent\cdot\tmsTrace}
    }{tms-trans}
  \end{center}
\end{definition}

\subsubsection{\gls*{sms} Monitor}\label{subsubsec-mon-sms}
placeholder
\subsubsection{\gls*{ms} Monitor}\label{subsubsec-mon-ms}
placeholder
\subsubsection{\gls*{scct} Monitor}\label{subsubsec-mon-scct}
placeholder
\subsubsection{\gls*{msscct} Monitor}\label{subsubsec-mon-msscct}
placeholder
\subsubsection{\gls*{ss} Monitor}\label{subsubsec-mon-spec}
placeholder
\subsubsection{\gls*{specms} Monitor}\label{subsubsec-mon-specms}
placeholder

>>>>

\paragraph{Spatial Memory Safety}
\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
For emphasis, we also use a custom trace-model for spatial memory safety.
\begin{definition}{\definitionlabel[\glsfirst{sms} Traces]{smstraces}}
\[
  \begin{array}{rcll}
    \text{(Base-Events)}&\smsEvent &:=& \smsAlloc \mid \smsUse \\
    \text{(Events)}&\smsEvent &:=& \smsEvent \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]
\end{definition}

@@

    \typerule{mapping-ms-sms-empty}{}{
      \mmap{\msev{\emptyevent}}{\smsev{\emptyevent}}
    }{mapping-ms-sms-empty}
    %
    \typerule{mapping-ms-sms-crash}{}{
      \mmap{\msev{\lightning}}{\smsev{\lightning}}
    }{mapping-ms-sms-crash}
    %
    \typerule{mapping-ms-sms-alloc}{}{
      \mmap{\msAlloc}{\smsAlloc}
    }{mapping-ms-sms-alloc}
    %
    \typerule{mapping-ms-sms-dealloc}{}{
      \mmap{\msDealloc}{\emptyevent}
    }{mapping-ms-sms-dealloc}
    %
    \typerule{mapping-ms-sms-use}{
      n=m
    }{
      \mmap{\msUse}{\smsUse}
    }{mapping-ms-sms-use}
  \end{center}
  \begin{center}
    \judgbox{\mmap{\msTrace}{\smsTrace}}{,,Translate \gls*{ms}-traces to \gls*{sms}-traces.''}

@@

    \typerule{mapping-ms-spec-empty}{}{
      \mmap{\msev{\emptyevent}}{\specev{\emptyevent}}
    }{mapping-ms-spec-empty}
    %
    \typerule{mapping-ms-spec-crash}{}{
      \mmap{\msev{\lightning}}{\specev{\lightning}}
    }{mapping-ms-spec-crash}
    %
    \typerule{mapping-ms-spec-alloc}{}{
      \mmap{\msAlloc}{\specAny[\unlock]}
    }{mapping-ms-spec-alloc}
    %
    \typerule{mapping-ms-spec-dealloc}{}{
      \mmap{\msDealloc}{\specAny[\unlock]}
    }{mapping-ms-spec-dealloc}
    %
    \typerule{mapping-ms-spec-use}{}{
      \mmap{\msUse}{\specAny[\unlock]}
    }{mapping-ms-spec-use}
  \end{center}
  \begin{center}
    \judgbox{\mmap{\msTrace}{\specTrace}}{,,Translate \gls*{ms}-traces to \gls*{ss}-traces.''}

@@

\subsubsection{\gls*{tms} Monitor}\label{subsubsec-mon-tms}

@@

    \typerule{tmsmon-refl}{
      \tmsMonitor[.A]=\emptyset
    }{
      \monitorcheck{\tmsMonitor}{\tmsMonitor}{\hole{\cdot}}
    }{tmsmon-refl}
    %
    \typerule{tmsmon-trans}{
      \monitorcheck[]{\tmsMonitor}{\tmsMonitor[']}{\tmsEvent} \rulesep
      \monitorcheck{\tmsMonitor[']}{\tmsMonitor['']}{\tmsTrace}
    }{
      \monitorcheck{\tmsMonitor}{\tmsMonitor['']}{\tmsEvent\cdot\tmsTrace}
    }{tms-trans}
  \end{center}
\end{definition}

\begin{definition}{\definitionlabel[Trace is \gls*{tms} via Monitor]{trace-is-tms-via-mon}}
  \begin{align*}
    \tmssafe\vdash\tmsTrace &\isdef \exists\tmsMonitor,\monitorcheck{\emptyset}{\tmsMonitor}{\tmsTrace}
  \end{align*}
\end{definition}

\begin{lemma}{\lemmalabel[Monitor \gls*{tms} is $\tmssafe$]{monitor-tms-is-tms}}
  If
  \begin{assumptions}
    \asm{monitor-tms-is-tms}{\tmssafe\vdash\tmsTrace}
  \end{assumptions}
  then
  \begin{goals}
    \goal{monitor-tms-is-tms}{\tmsTrace\in\tmssafe}
  \end{goals}
\end{lemma}
\begin{proof}
  \incompleteProof
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{\gls*{sms} Monitor}\label{subsubsec-mon-sms}
\begin{definition}{\definitionlabel[\gls*{sms} Monitor]{monitor-sms}}
  \begin{gather*}
    \begin{aligned}
      \text{(Abstract Store)}~\smsMonitor &\isdef \mi{Locs}\times\mathbb{N} \\
      \operatorname{dom}\lrpars{\smsMonitor}&=\{\varLoc \mid \forall n, \lrpars{\varLoc;n}\in\smsMonitor\}
    \end{aligned}
  \end{gather*}
\end{definition}

The reductions are defined as follows:
\begin{definition}{\definitionlabel[\gls*{sms} Monitor Reductions]{monitor-sms-steps}}
  \begin{center}
    \judgbox{\monitorcheck[]{\smsMonitor}{\smsMonitor[']}{\smsEvent}}{,,Given abstract state $\smsMonitor$, step to abstract state $\smsMonitor'$\\ emitting an abstract event $\smsEvent$.''}

    \typerule{smsmon-empty}{}{
      \monitorcheck[]{\smsMonitor}{\smsMonitor}{\emptyevent}
    }{smsmon-empty}
    \typerule{smsmon-alloc}{
      \varLoc\not\in\operatorname{dom}\lrpars{\smsMonitor}
    }{
      \monitorcheck[]{\smsMonitor}{\{\varLoc\}\cup\smsMonitor}{\smsAlloc}
    }{smsmon-alloc}
    \typerule{smsmon-use}{
      \lrpars{\varLoc;n}\in\smsMonitor \rulesep
      m<n
    }{
      \monitorcheck[]{\smsMonitor}{\smsMonitor}{\smsUse}
    }{smsmon-use}
    \typerule{smsmon-crash}{}{
      \monitorcheck[]{\smsMonitor}{\smsMonitor}{\lightning}
    }{smsmon-crash}
  \end{center}
  \begin{center}
    \judgbox{\monitorcheck{\smsMonitor}{\smsMonitor[']}{\smsEvent}}{,,Reflexive-Transitive closure of above judgement.\\$\smsTrace$ is a list of events.''}

    \typerule{smsmon-refl}{
      \smsMonitor=\emptyset
    }{
      \monitorcheck{\smsMonitor}{\smsMonitor}{\hole{\cdot}}
    }{smsmon-refl}
    %
    \typerule{smsmon-trans}{
      \monitorcheck[]{\smsMonitor}{\smsMonitor[']}{\smsEvent} \rulesep
      \monitorcheck{\smsMonitor[']}{\smsMonitor['']}{\smsTrace}
    }{
      \monitorcheck{\smsMonitor}{\smsMonitor['']}{\smsEvent\cdot\smsTrace}
    }{smsmon-trans}
  \end{center}
\end{definition}

\begin{definition}{\definitionlabel[Trace is \gls*{sms} via Monitor]{trace-is-sms-via-mon}}
  \begin{align*}
    \smssafe\vdash\smsTrace &\isdef \exists\smsMonitor,\monitorcheck{\emptyset}{\smsMonitor}{\smsTrace}
  \end{align*}
\end{definition}

\begin{lemma}{\lemmalabel[Monitor \gls*{sms} is $\smssafe$]{monitor-sms-is-sms}}
  If
  \begin{assumptions}
    \asm{monitor-sms-is-sms}{\smssafe\vdash\smsTrace}
  \end{assumptions}
  then
  \begin{goals}
    \goal{monitor-sms-is-sms}{\smsTrace\in\smssafe}
  \end{goals}
\end{lemma}
\begin{proof}
  \incompleteProof
\end{proof}

\subsubsection{Dual-Monitor}\label{subsubsec-mon-dual}

\begingroup
  \newcommand{\aM}[1][]{\varMonitor[A]#1}
  \newcommand{\bM}[1][]{\varMonitor[B]#1}
  \newcommand{\aE}[1][]{\varEvent[A]#1}
  \newcommand{\bE}[1][]{\varEvent[B]#1}

For sake of this section, let $A,B$ be monitors with respective monitor states $\aM$ and $\bM$.
Then, both monitor states can be combined as follows:

\begin{definition}{\definitionlabel[Dual-Monitor State]{monitor-dual}}
  \begin{gather*}
    \begin{aligned}
      \text{(Abstract State)}~\dualMonitor &\isdef \aM \times \bM\\
    \end{aligned}
  \end{gather*}
\end{definition}

The reductions are defined as follows:
\begin{definition}{\definitionlabel[Dual-Monitor Reductions]{monitor-dual-steps}}
  \begin{center}
    \typerule{dualmon-refl}{}{
      \monitorcheck{\lrpars{\emptyset_A;\emptyset_B}}{\lrpars{\emptyset_A;\emptyset_B}}{\hole{\cdot}}
    }{dualmon-refl}
    %
    \typerule{dualmon-trans-emptyA}{
      \dualMonitor=\lrpars{\aM;\bM} \rulesep
      \monitorcheck[]{\bM}{\bM}{\bE} \rulesep
      \monitorcheck{\dualMonitor}{\dualMonitor[']}{\varTrace[_{A;B}]}
    }{
      \monitorcheck{\dualMonitor}{\dualMonitor[']}{\lrpars{\emptyevent;\bE}\cdot\varTrace[_{A;B}]}
    }{dualmon-trans-emptyA}
    %
    \typerule{dualmon-trans-emptyB}{
      \dualMonitor=\lrpars{\aM;\bM} \rulesep
      \aE\not=\emptyevent \rulesep
      \monitorcheck[]{\aM}{\aM}{\aE} \rulesep
      \monitorcheck{\dualMonitor}{\dualMonitor[']}{\varTrace[_{A;B}]}
    }{
      \monitorcheck{\dualMonitor}{\dualMonitor}{\lrpars{\aE;\emptyevent}\cdot\varTrace[_{A;B}]}
    }{dualmon-trans-emptyB}
    %
    \typerule{dualmon-trans}{
      \lrpars{\aE;\bE}\not=\lrpars{\emptyevent;\emptyevent} \rulesep
      \monitorcheck[]{\aM}{\aM[']}{\aE} \rulesep
      \monitorcheck[]{\bM}{\bM[']}{\bE} \rulesep
      \monitorcheck{\dualMonitor}{\dualMonitor[']}{\varTrace[_{A;B}]}
    }{
      \monitorcheck{\dualMonitor}{\dualMonitor}{\lrpars{\aE;\bE}\cdot\varTrace[_{A;B}]}
    }{dualmon-trans}
  \end{center}
\end{definition}

\begin{definition}{\definitionlabel[Dual-Trace Satisfaction via Monitor]{trace-is-ab-via-mon}}
  Let $A,B$ monitors and $\varProperty$ a property,
  \begin{align*}
    \varProperty\vdash\varTrace[A;B] &\isdef \exists\aM\;\bM,\monitorcheck{\emptyset}{\lrpars{\aM;\bM}}{\varTrace[A;B]}
  \end{align*}
\end{definition}

% \begin{lemma}{\lemmalabel[Monitor \gls*{sms} is $\smssafe$]{monitor-sms-is-sms}}
%   If
%   \begin{assumptions}
%     \asm{monitor-sms-is-sms}{\smssafe\vdash\smsTrace}
%   \end{assumptions}
%   then
%   \begin{goals}
%     \goal{monitor-sms-is-sms}{\smsTrace\in\smssafe}
%   \end{goals}
% \end{lemma}
% \begin{proof}
%   \incompleteProof
% \end{proof}
\endgroup

\subsubsection{\gls*{ms} Monitor}\label{subsubsec-mon-ms}
Instantiate \Cref{subsubsec-mon-dual} with \Cref{subsubsec-mon-tms,subsubsec-mon-sms}. 
What is left to give is a translation from traces $\msTrace$ to traces $\varTrace[\tmssafe;\smssafe]$.
This translation is done as follows:
\begin{definition}{\definitionlabel[\gls*{ms} to \gls*{tms} and \gls*{sms}]{traces-ms-to-tmssms}}
  ...
\end{definition}

\subsubsection{\gls*{scct} Monitor}\label{subsubsec-mon-scct}
placeholder
\subsubsection{\gls*{msscct} Monitor}\label{subsubsec-mon-msscct}
placeholder
\subsubsection{\gls*{ss} Monitor}\label{subsubsec-mon-spec}
placeholder
\subsubsection{\gls*{specms} Monitor}\label{subsubsec-mon-specms}
placeholder
